---
title: "Aides Drug"
author: "Giselle Gonzalez"
date: "Summer 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)

druggies = select(triUmp, c(treatment, infected, DrugUser))
druggies = filter(druggies, DrugUser == "yes")


```


# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$:The drug is not effective to drugusers  
$H_A$:The drug is effective to drugusers

# Methods

 treatment is a categorical variable  
 infected is a numerical  
 so i use the cat~num helper file
 
 
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

 
## Descriptive Results

### Graphical Descriptive Results

```{r}
barchartGC(~infected + treatment,data=druggies)
barchartGC(~infected + treatment,data=druggies, type="percent")

```
 
 
 This shows that the people who took the placebo had more of a reaction to the drug users who didnt take the treatment 
 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->



### Numerical Descriptive Results


```{r}
table1 <- xtabs(~infected + treatment,data= druggies)
rowPerc(table1)
colPerc(table1)
```

This data shows that the people who to the placebo had 75 percent moreof a reaction to the treatment  than the drug users who didnt had a 24 percent

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi square p-value is 0.0000001 which is lower than 0.05  
The fishers p-value is 0.00000004 it is also lower than 0.05 so we rejct the null  
The odds ratio is 0.4:1 meaning that the drug users who took the placebo are 0.4 less likely to be effective than the 1 who didnt 


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
